Statistical analysis of periodic data in neuroscience

### 2021-01-12 manuscript/manuscript.Rmd
Preparing for first public release

nsims <- 100  # global value for the whole script - increase to 100000 for final run, but note this takes around 16 hours to complete (depending on processor speed)

@@

  polygon(0.4*ellRot2[1, ]+xlocs[m], 0.1*ellRot2[2, ]+1.5, col=addalpha(nbdtpal[2],0.3),border=NA)
  lines(0.4*ellRot[1, ]+xlocs[m], 0.1*ellRot[2, ]+1.5, lwd=1)
  lines(0.4*c(ctr2[1],xdata[1])+xlocs[m],0.1*c(ctr2[2],ydata[1])+1.5,col=nbdtpal[2],lwd=3)
  points(0.4*xdata+xlocs[m],0.1*ydata+1.5,pch=16,cex=0.5,col=nbdtpal[3])
  points(0.4*ctr2[1]+xlocs[m],0.1*ctr2[2]+1.5,pch=16,cex=1,col=nbdtpal[1])
  points(0.4*xdata[1]+xlocs[m],0.1*ydata[1]+1.5,pch=16,cex=1,col=nbdtpal[2])
}

@@

nsig <- matrix(0,nrow=length(samplesizes),ncol=length(dlist))
for (s in 1:length(samplesizes)){
  for (d in 1:length(dlist)){
    for (n in 1:nsims){
    data <- matrix(rnorm(2*samplesizes[s]),nrow=samplesizes[s],ncol=2)
    data[1,] <- c(dlist[d],0)
    output <- CI.test(data)
    if(output$pval<0.05){nsig[s,d] <- nsig[s,d] + 1}
    }
  }
}

@@

for (s in 1:length(samplesizes)){
  lines(dlist,propsig[s,],lwd=3,col=nbdtpal[s])
}

>>>>

nsims <- 100000  # global value for the whole script - increase to 100000 for final run, but note this takes around 16 hours to complete (depending on processor speed)

@@

  polygon(0.4*ellRot2[1, ]+xlocs[m], 0.1*ellRot2[2, ]+1.5, col=addalpha(nbdtpal[2],0.3), border=NA)
  lines(0.4*ellRot[1, ]+xlocs[m], 0.1*ellRot[2, ]+1.5, lwd=1)
  lines(0.4*c(ctr2[1],xdata[1])+xlocs[m],0.1*c(ctr2[2],ydata[1])+1.5,col=nbdtpal[2],lwd=3)
  points(0.4*xdata+xlocs[m],0.1*ydata+1.5,pch=16,cex=0.5,col=nbdtpal[3])
  points(0.4*ctr2[1]+xlocs[m],0.1*ctr2[2]+1.5,pch=16,cex=1,col=nbdtpal[1])
  points(0.4*xdata[1]+xlocs[m],0.1*ydata[1]+1.5,pch=16,cex=1,col=nbdtpal[2])
}

@@

# this version avoids calling the condition index test to speed things up
# it means that the probability density function is not calculated on each iteration
cilist <- seq(1,100,0.001)    # list of possible CIs
alpha <- 0.05
nsig <- matrix(0,nrow=length(samplesizes),ncol=length(dlist))
for (s in 1:length(samplesizes)){

  N <- samplesizes[s]
  # calculate probability density function using modified Edelman equation
  pdffunction <- ((N-2)*(2^(N-2))) *
    ((cilist^2 - 1)/((cilist^2+1)^(N-1))) * (cilist^(N-3))
  cdfinverse <- 1-(cumsum(pdffunction)/sum(pdffunction)) # inverse of cdf
  criticalCI <- cilist[min(which(cdfinverse<=alpha))]  # find the threshold CI

  for (d in 1:length(dlist)){
    for (n in 1:nsims){
      data <- matrix(rnorm(2*samplesizes[s]),nrow=samplesizes[s],ncol=2)
      data[1,] <- c(dlist[d],0)

      C <- cov(data)  # covariance matrix
      N <- nrow(data) # sample size
      eigVal  <- eigen(C)$values  # calculate eigenvalues
      CI <- sqrt(eigVal[1]/eigVal[2]) # calculate condition index

      if(CI>criticalCI){nsig[s,d] <- nsig[s,d] + 1}
    }
  }
}

@@

for (s in 1:length(samplesizes)){lines(dlist,propsig[s,],lwd=3,col=nbdtpal[s])}
